# Service Design and Business Case Development {#service}

We decided to take a new, modern approach to the ‘data observatory’ concept, and modernize it with the application of 21st century data and metadata standards, the new results of reproducible research and data science. Various UN and OECD bodies, and particularly the European Union support or maintain more than 60 data observatories, or permanent data collection and dissemination points, but even these do not use these organizations and their members open data. 

We are building open-source data observatories, which run open-source statistical software that automatically processes and documents reusable public sector data (from public transport, meteorology, tax offices, taxpayer funded satellite systems, etc.) and reusable scientific data (from EU taxpayer funded research) into new, high quality statistical indicators.

We are building various open-source data collection tools in R and Python to bring up data from big data APIs and legally open, but not public, and not well served data sources. For example, we are working on capturing representative data from the Spotify API or creating harmonized datasets from the Eurobarometer and Afrobarometer survey programs.

```{r value-added, include=FALSE, fig.align="Service Flow"}
knitr::include_graphics(
  here::here("plots", "slides", "automated_observatory_value_chain.jpg")
)
```


Open data is usually not public; whatever is legally accessible is usually not ready to use for commercial or scientific purposes. In Europe, almost all taxpayer funded data is legally open for reuse, but it is usually stored in heterogenous formats, processed into an original government or scientific need, and with various and low documentation standards. Our expert data curators are looking for new data sources that should be (re-) processed and re-documented to be usable for a wider community.

We would like to introduce our service flow, which touches upon many important aspects of data scientist, data engineer and data curatorial work.

- We believe that even such generally trusted data sources as Eurostat often need to be reprocessed, because various legal and political constraints do not allow the common European statistical services to provide optimal quality data – for example, on the regional and city levels.
- With rOpenGov and other partners we are creating open-source statistical software in R to re-process these heterogenous and low-quality data into tidy statistical indicators, and automatically validate, and document it.
- We are carefully documenting and releasing administrative, processing, and descriptive metadata, following international metadata standards, to make our data easy to find and easy to use for data analysist.
- We are automatically creating depositions and authoritative copies marked with an individual digital object identifier (DOI) to maintain data integrity.
- We are building a simple database and supporting API that releases the data without restrictions, in a tidy format that is easy to join with other data, or easy to join into databases, together with standardized metadata.
- We maintain observatory website where not only the data is available, but we are placing tutorials and use cases to make it easier to use them. Our aim is to show a modern, 21st century reimagination of the data observatory concept developed and supported by the UN, EU and OECD, and we want to show that modern reproducible research and open data could make the existing 60 data observatories and the planned new ones grow faster into data ecosystems.

We are working around the open collaboration concept, which is well-known in open source software development and reproducible science, but we try to make this agile project management methodology more inclusive, and include data curators, and various institutional partners into this approach. Based around our early-stage startup, Reprex, and the open-source developer community rOpenGov, we are working together with other developers, data scientists, and domain specific data experts in climate change and mitigation, antitrust and innovation policies, and various aspect of the music and film industry.

```{r, fig.cap="Our open collaboration is truly open: new data curators, data scientists and data engineers are welcome to join in."}
knitr::include_graphics(
  here::here("plots", "screenshots", "dmo_contributors.png")
)
```

Our open collaboration is truly open: new data curators, data scientists and data engineers are welcome to join in.  We develop open-source software in an agile way, so you can join in with an intermediate programming skill to build unit tests or add new functionality, and if you are beginner, you can start with documentation and testing our tutorials.  For business, policy, and scientific data analysts, we provide unexploited, exciting new datasets.  And of course, advanced developers can join our development team: the statistical data creation is mainly made in the R language, and the service infrastructure in Python and Go components.

## Professional Standards in Business {#business-professional-standards} 

### Key Business Indicators


### Business Record Retention 

Some of the requirements of reproducible research are usually required by professional standards.  For example, various accounting, finance, legal or consulting professional standards call for appropriate documentation and record retention.

## Professional Standards in Policy {#policy-professional-standards} 

### Evidence-based, Open Policy Analysis {#opa}

In the last two decades, governments and researchers have placed a growing emphasis on the value of evidence-based policy. However, while the evidence generated through research to inform policy has become more rigorous and transparent, policy analysis--the process of contextualizing evidence to inform specific policy decisions--remains opaque.  

We believe that a modern data observatory must improve how evidence is created and used in policy reports, and pass on the efficiency gains from increasing reproducibility and automation. Therefore, we pledge that the [music.dataobservatory.eu](https://music.dataobservatory.eu) will comply with the [Open Policy Analysis](https://www.bitss.org/opa/) standards developed by the [Berkeley Initiative for Transparency in the Social Sciences](https://www.bitss.org/) & [Center for Effective Global Action](https://cega.berkeley.edu/). These standards are applied by the World Bank.

## Contributors

We are looking for business associates to help our service design into our

- [Green Deal Data Observatory](https://greendeal.dataobservatory.eu/#contributors)

- [Economy Data Observatory](https://economy.dataobservatory.eu/#contributors)

- [Digital Music Observatory](https://music.dataobservatory.eu/#contributors) 

1. Work with governmental or scientific or otherwise [open data](#open-data).

2. Committed to high policy or business professional standards, and by making their work [reproducible](#reproducible-research), they adhere to reviewability, reproducability, confirmability and auditability, regardless if they work, or study for various professsional roles in business, academia, public or non-governmental policy, and data journalism.

>An important aspect of the EU Datathon Challenges is ".. to propose the development of an application that links and uses open datasets [...] to find suitable new approaches and solutions to help Europe achieve important goals set by the European Commission through the use of open data."

Where to find us:
- [dataobservatory-eu](https://github.com/dataobservatory-eu) is our private repo collection and private github collaboration platform, but many of our repos are open. Like this one.

- [Creative Data Observatories LinkedIn Page](https://www.linkedin.com/company/reprexbv/). Make sure you follow us, and spread our messages.

- [twitter.com/dataandlyrics](https://twitter.com/dataandlyrics) is our twitter handle for our music-oriented blog. If you are on twitter, please follow us, and retweet our blogposts.

- [keybase.io/team/reprexcommunity](https://keybase.io/team/reprexcommunity) is our landing page to our otherwise private and invisible internal communication platform. (See more in subchapter\@ref(keybase) of [Tools](#keybase).)

You find more information in the \@ref(data-curators) chapter about our topics: [What is Open Data?](#open-data), [Reproducible Research](#reproducible-research-theory), and of course, [Get Inspired About Data](#get-inspired).

## Passion About Our Topic {#topic-associate}

- You are passionate about one of our topics, but you do not feel that you have the skills yet to become a data curator or a developer. 

- You have a curiosity in the field of economic policies, particularly in computational antitrust, innovation research, and understanding the statistically under-represented micro- and small enterprises, join our [Economy Data Observatory](https://economy.dataobservatory.eu/#contributors) as a volunteer.

- You are passionate about environmental research of climate change, designing urban, social and economic mitigation strategies, or undertanding how people think about climate change, join our [Green Deal Data Observatory](https://greendeal.dataobservatory.eu/#contributors)  team as a volunteer.  

- You want to know how musicians can make a living after the pandemic?  How can we make sure that music made by womxn, small country artists or artists of color gets and equal chance? Are you interested in the future of ethical AI and data governance? Join our [Digital Music Observatory](https://music.dataobservatory.eu/#contributors) team as a volunteer.

## Passion For Trustworthy AI, Open Data and Open Source Projects {#open-associate}

- You want to learn to write scientifically valid, open source code in R or Python, but you are a beginner. We help you anywhere - you can even start copyediting or technical documentation (it is a must in open source development) and create tutorials for you to interact with our data and products (if if helps you, it will help others.)

- As a business economist or legal professional, you are interested how open data, open source software, research automation and ethical, trustworthy AI products can find their market. 

- As a blogger, data journalist or marketeer you would like to help to make open data, and transparent, ethical, open AI more widely known and used.

## Technical Requirements {#technical-requirements-associate}

- Make sure that you read the [Contributors Covenant](https://www.contributor-covenant.org/). You must make this [pledge](https://www.contributor-covenant.org/version/2/0/code_of_conduct/) to make participation in our community a harassment-free experience for everyone, regardless of age, body size, visible or invisible disability, ethnicity, sex characteristics, gender identity and expression, level of experience, education, socio-economic status, nationality, personal appearance, race, caste, color, religion, or sexual identity and orientation. Participating in our data observatories requires everybody to act and interact in ways that contribute to an open, welcoming, diverse, inclusive, and healthy community. It's better this way for you and for us!

- Give users at least one social media account where they can get in touch with you (any of LinkedIn, Twitter, Academia, SSRN, Google Scholar, or even Facebook.)

- Please, follow us on social media, it really helps us finding new users and showing that we are able to grow our ecosystem.
  - [Green Deal Data Observatory on Linkedin](https://www.linkedin.com/company/78556699) and [Green Deal Data Observatory on Twitter](https://twitter.com/GreenDealObs)
  - [Economy Data Observatory on Linkedin](https://www.linkedin.com/company/78562153) and [Economy Data Observatory on Twitter](https://twitter.com/GreenDealObs)
  - [Digital Music Data Observatory on Linkedin](https://www.linkedin.com/company/reprexbv/) and [Digital Music Data Observatory on Twitter](https://twitter.com/dataandlyrics)

- Please send us back this [md file](https://greendeal.dataobservatory.eu/media/documents/bio_template.md) with your data. You can open it with any text editor, but Notepad, TextEdit, Vim and similar clean text editors are the best. 

- If you feel you need chatting on onboarding, contact us on [Keybase](https://curators.dataobservatory.eu/tools.html#keybase) - it's lightweight, discrete, encrypted, your mother, partner and friends are not there, it is free, open source, and can share/exchange files, too. Otherwise in email.
