---
title: "Blogpost"
author: "Daniel Antal, CFA"
date: "6/16/2021"
- book.bib
- antal.bib
- dcms.bib
- valuation.bib
- ccipolicy.bib
- contentregulation.bib
- caselaw.bib
- eulaw.bib
- musicindustry.bib
- musicology.bib
- packages.bib
- statsoftware.bib
- statisticalmethodology.bib
- youtube.bib
- musiceducation.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reprex signed an agreement with two investigators of two principal investigators of the [Music Creators’ Earnings in the Streaming Era](https://digit-research.org/research/related-projects/music-creators-earnings-in-the-streaming-era/), project, [Prof David Hesmondhalgh](https://music.dataobservatory.eu/author/prof-david-hesmondhalgh/) and [Dr Hyojugn Sun](https://music.dataobservatory.eu/author/hyojung-sun/), to provide data for their research report in an accompanying, reproducible research document, to deposit the data of the project in a reproducible manner in the Digital Music Observatory, and to promote the use of the Digital Music Observatory. The project is funded by the Research England QR Special Projects Fund and the UK Intellectual Property Office.

The music industry requires a permanent market monitoring facility to win fights in competition tribunals, because it is increasingly disputing revenues with the world's biggest data owners. 
This was precisely the role of the former CEEMID [@artisjus_measuring_2014] program, that was initiated by a few collective management societies after a dropped GESAC project. 

Starting out from three relatively data-poor countries, where data pooling allowed rightsholders to increase revenues, the CEEMID data collection program was extended by 2019 to 12 countries.  It was eventually transformed to the [Demo Music Observatory](https://music.dataobservatory.eu/) in 2020 [@antal_launching_music_observatory_2021], which is now open for any national rightsholder, stakeholder organization or music research institute. 

The idea of this observatory was brought to the UK policy debate on music streaming by the observatory's only (former) British users, via the _Written evidence submitted by The state51 Music Group_ to the _Economics of music streaming review_ of the DCMS Committee [@dcms_state51_2020]. The state51 music group was kind not to limit the use of the earlier published dataset for this project.

```{r}
plot(stream_example_release_all_plot)
```

Stating that the greatest difference among rightsholders' earnings is related to the popularity of their works and recorded fixations appears banal and trivial at first. Yet, because many payout problems appear in the hard-to-describe long tail, understanding the _justified_ differences of rightsholder earnings is a very important step towards identifying the _unjustified_ differences. The principle of equitable remuneration, enshrined in the international copyright system and UK law, makes it compulsory to reflect on differences in popularity, or in other words, on the use of the work or its fixation in royalty distributions. It would be a breach of copyright law if less popular, or never played artists, would receive significantly more payment at the expense of popular artists. The earnings must reflect the difference in use and the economic value in use among rightsholders. 

## Justified Difference & Changes Over Time

Among the _justified_ differences we quantify four objective justifications: 

1. *The variability of the domestic price of a stream over time* in the subchapter entitled *Effect of Release Time*. We observe a diminishing, but variable value of streams. Depending on the release date of a recording, and how quickly it builds up or loses the interest of the audience, the same number of streams can result in about 28% different earnings.

2. *The variability of international market share* and international streaming prices in *International Competitiveness*. Compared to domestic streaming prices, most international markets, particularly emerging markets, have a much greater variability of streaming prices. Artists who have a significant foreign presence, depending on their foreign market share, can experience 2-3 times greater differences in earnings than artists whose audience is predominantly British. The variability of prices in advanced foreign markets such as Germany was similar to that of the British market, but in  emerging markets and smaller advanced markets--such as the Netherlands--we measured a variability of around 50-80%. 

3. *The variability of the exchange rate* that is applied when translating foreign currency revenues to the British pound in *Exchange Rate Effects* Our [CEEMID-CI Streaming Indexes](#ceemid-ci) index covers the post-Brexit referendum period, when the British pound was generally depreciating against most currencies. This resulted in a GBP-denominated translation gain for artists with a foreign presence. We show that the variability of the GBP exchange rates can add _bigger justified differences_ among rightsholders' earnings than the entire British price variation. The exchange rate movements are typically in the range of 30%, or at the level of the British domestic price variations in streaming prices. In our simulated results, this effect shielded the internationally competitive rightsholders from a significant part of the otherwise negative price change in foreign markets. 

4. The choice of *distribution model*, i.e. *pro-rata* or *user-centric* distribution (see *Distortion of Earnings Distributions*. We did not have access to British data to correctly understand the distribution effects in time and model a possible change from the *pro-rata* to the *user-centric* system. Based on foreign-country research, we show that a zero-sum distribution change could potentially increase the income of less popular and older artists at the expense of very popular and younger artists. We believe that a typical redistribution for less popular, long-tail artists would add typically 15% at the expense of the most popular artists. 

We clearly believe that the distribution choice between the *pro-rata* and *user-centric* model is still within the scope of _justified_ differences. In our understanding, both models are in line with the legal requirements of _equitable remuneration_, and the choice between the two models pits various rightsholder groups against each other in a zero-sum game. We instead propose an alternative approach of *artist-centric distribution* that could be potentially a win-win for all rightsholder groups; further elaboration of the concept lies outside the scope of this report. 

## Unjustified Differences


In our understanding, there are some known and some hypothetical causes of _unjustified_ earning differences.

1. We did not have systemic data on the *uncollected revenues* (see \@ref(uncollected-simulation) [Uncollected Revenues](#uncollected-simulation) -- these are earnings that are legally made, but due to documentation, matching, processing, accounting, or other problems, the earnings are not paid. We could not even attempt to estimate this problem in the absence of relevant British empirical data. In our subjective opinion, based on consultations with various organizations managing royalty payments, the problem is likely to be greater in the case of composers than in producer and performer revenues. In ideal cases, of course, the unclaimed royalty is near 0% of the earnings; it seems that on advanced markets the magnitude of this problem is in the single digits, and in emerging markets much greater, sometimes up to 50%. Royalty distribution is a costly business, and the smaller the revenue, the smaller the cost base to manage billions of transactions and related micropayments.

2. More analysis is required to understand how the algorithmic, highly autonomous recommendation systems of digital platforms such as Spotify, YouTube, Apple, and Deezer, among others, impact music rightsholders' earnings. There are many research projects in the UK and Europe that aim at making these systems accountable and trustworthy. Currently, we do not know if the algorithms are biased in terms of gender, sexuality, age, race, ethnicity, ability, or genre, among others. Our Demo Music Observatory aims to collect large enough observations to make advanced casual analysis in this field. There are some empirical findings that suggest that such biases are present in various platforms, but due to the high complexity of recommendation systems, it is impossible to intuitively assign blame to pre-existing user biases, wrong training datasets, improper algorithm design, and other factors (see \@ref(trustworthy-ai) [Trustworhty AI Problems](#trustworthy-ai)).

3. There is always a hypothetical possibility that organizations with monopolistic power try to corner the market or make the playing field uneven. The music industry requires a permanent market monitoring facility to win fights in competition tribunals, because it is increasingly disputing revenues with the world's biggest data owners. (See \@ref(observatory)[Ongoing Data Collection and Market Monitoring: The Data Observatory Concept](#observatory).)

## Level of Earnings
For example, the current debate in the United Kingdom is often thematised around the submission of the *#BrokenRecord* campaign to the DCMS committee, which calls for a legal re-definition of such rights [@dcms_brokenrecord_2020]. This idea is not unique to the United Kingdom, in various European jurisdictions, performers fought similar campaigns for legislation or went to court, sometimes successfully, for example, in Hungary. While the general concept of *equitable remuneration* is the same in the United Kingdom, Spain or Hungary, the way the rights are constructed and exercised are very different in different legal systems. So generally one cannot make the assumption that if one jurisdiction gives this right to one group of rightsholders, they can have the exact same right in another.

The re-definition of copyrights and neighboring rights, or, as a fact, any intellectual property, is not unusual. These are legal monopolies, which are given to encourage artistic and innovative creation, and they are subject to *ad hoc* reviews, as the technologies and the social and economic context is changing for exploiting these legal monopolies in the form of intellectual property rights, such as copyrights or neighboring rights. 

The current streaming licensing model in Europe is a mix of mechanical and public performance rights. Therefore, when we are talking about music streaming, we must compare the streaming sub-market with the digital downloads, physical sales and private copying markets (mechanical licensing), and with the radio, television, cable and satellite retransmission markets (public performance licensing.) 


At last, we would like to draw attention to the _level of earnings_.  In the authors' international experience, the British music industry is not the only one where rightsholders are too much focused on dividing up their revenues among each other, instead of increasing the size of the pie. As we show in our forthcoming report, the legal concept *equitable remuneration* is inseparable from the economic concept of *fair valuation*. The concept of _fair value_ in music streaming, which is a legal claim of all British rightsholders, is impossible to establish by studying *only* the economics of music streaming. Music streaming is not a separate market activity. Streaming services are competing with digital downloads, physical sales, and radio broadcasting; and the media streaming of YouTube and similar services is competing with music streaming, radio, and television broadcasting as well as retransmissions. Because music streaming, and the media streaming model of YouTube and other platforms is gaining share against digital downloads and private copying (mechanical royalty revenues) and against various forms of broadcast or retransmitted radio and television (public performance revenues), it is a critically important question if the new streaming licensing model (which is in itself a mix of the mechanical and public performance licensing) is capable of replacing the revenues for all rightsholders from the earlier services and sales channels that it is replacing.

The replacement of various mechanical and public performance revenues with streaming revenues affected various rightsholder groups very differently.  It is perfectly plausible, that an important group of producers, supported by their early advantage in algorithmic sales to gain foreign market shares, and supported by a devaluing British pound, did experience significant revenue growth in their own enterprises in the last 5-10 years, while other producers, and particularly publishers or performers experienced a stark decline. Our analysis in other countries, that we could not replicate in the United Kingdom, shows that the streaming model replaces only a fraction of the earlier total composer, producer and performer revenues \@ref(equitable-historical) [Comparison with Mechanical and Historical Revenues](#equitable-historical).  

We made very detailed calculations in four years in Hungary comparing the total revenues from mechanical, public performance and mixed streaming licensing models, and our experience was that licensed streaming (Apple, Deezer, Spotify) was replacing revenues on the level of public performance revenues, but far lower than earlier mechanical revenues, which resulted in the devaluation of music. This loss was not equally shared by the three rightsholders groups. We measured the monetary value of the _value gap_, too.  Media platforms (like YouTube or Facebook) use music differently than licensed music streaming platforms. They replace mechanical and public performance revenues, and even private copying compensation, but at a far lower rate in our experience \@ref(full-market) [Comparison with Other Revenues](#full-market). The European Union decided to try to close the _value gap_ with the new copyright directive, but all not yet conclusive evidence points to the direction that the chosen method will not bring platform revenues to the level of public performance, mechanical or mixed music streaming licensing. In the United Kingdom, the choice of not applying the DCM Directive opens the door to try to more successfully recover the devaluation on this sales channel.

The _equitable remuneration_ and _fair level_ of music streaming earnings cannot be subject to a valid economic analysis without analyzing *the economics of music*. Music streaming, and media platform streaming are competing with the traditional mechanical and public performance licensing challenge, and if they are not able to replace the revenues earned on the replaced volumes (in enjoyment hours or other metrics) then music will devalue.  This devaluation is different among various rightsholder groups -- small groups may even gain, while the majority will lose out. If the devaluation reaches a critical level, rightsholders and the intellectual property policy maker have two choices: try to change the outcome in courts, particularly in copyright or competition tribunals, or redefine the intellectual property right with a higher expected value. Unfortunately for the music industry, while competition law can be a potent legal force, it is applied mainly against too high, and not too low prices. 

In other countries, I was able to analyze the entire market--not only its media and music streaming segments--and to make representative surveys of rightsholders' revenues. In these countries, in the same period, we could observe a general devaluation of music. It is very unlikely that the United Kingdom is exempt from these international trends, because as we show in the \@ref(#empirical). chapter on [Empirical Earnings](#empirical), the streaming prices, as measured by the [CEEMID-CI value and volume](#streaming-price) indexes, are declining in the United Kingdom, too. My experience in other countries, partly backed by harmonized data from the United Kingdom, shows that the last ten years were characterized by two large trends: a decline in unpaid, uncompensated illegal use of music, and a substitution of higher value licenses to lower value licenses -- and in the case of media platforms like YouTube, to a lower value and new remuneration method (that is often associated with the value gap or transfer of value.) The overall effect in the countries I have analyzed in detail was negative. Streaming not only competes with uncompensated illegal use, but replaces licenses that often yielded much higher revenues. In addition, media platforms have moved from and advertising revenue base behind public performance royalties to an advertisement-sharing system that has resulted in far lower revenues.

It would be irresponsible to simply extrapolate from European market experience. However, some of the analysis and models conducted by European rightsholders in the last years would be very easy to replicate in the UK, and it is unlikely to yield different results. The very idea of our Music Observatory, which was placed into the UK government's Department for Digital, Culture, Media and Sport (DCMS) inquiry by Consolidated Independent, is based on the idea that the economics of music, and the deteriorating compensation of rightsholders can only be understood by joining many data sources and analyzing the inherently international economics of music based on international data sharing. 

Music rightsholders' earnings could only rise if systematic changes were made by the industries, for example, with fixing the so-called value transfer, or identifying underperforming licenses and renegotiating them. A change to the user-centric distribution method, for instance, would potentially undermine the music industry's position, as we explain in \@ref(earnings-distorsion). The industry should be better off if the energies focused on the zero-sum game of the *pro-rata* versus *user-centric* payments would be put into finding a new pricing distribution model for music streaming that has a positive expected income.


## Solidarity
Music earnings are very heavily skewed towards a small number of very successful composers, performers, and producers. Compared to some other licensing forms, the streaming model has little elements of solidarity *Solidarity Equitable Remuneration and more Equal Distribution*. While many licensing models--particularly public performance licensing and in some countries, private copying remuneration--have strong solidaristic elements for the financially less successful, or for music deemed to be of a specifically designated cultural import, streaming appears to be working in favor of successful artists and the larger labels to the detriment of smaller actors. However, in those cases, the solidarity element is decided by the rightsholders themselves, and not by external parties like radio broadcasters or streaming service providers. The so-called socio-cultural funds that provide assistance for artists in financial need must be managed by rightsholders, not others, and the current streaming model makes the organizations of such solidaristic action particularly difficult.
