--- 
title: "Automated Observatory Contributors' Handbook"
author: "Daniel Antal, CFA"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
lof: yes
classoption: openany, a4paper, oneside
papersize: a4
geometry: left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm
bibliography:
- book.bib
- antal.bib
- dcms.bib
- valuation.bib
- ccipolicy.bib
- contentregulation.bib
- caselaw.bib
- eulaw.bib
- musicindustry.bib
- musicology.bib
- packages.bib
- statsoftware.bib
- statisticalmethodology.bib
- youtube.bib
- musiceducation.bib
biblio-style: apalike
linestretch: 1.1
urlcolor: blue
linkcolor: blue
link-citations: yes
description: "We want to win an EU Datathon prize by processing the vast, already available governmental and scientific open data made usable for policy-makers, scientific researchers, and business researcher end-users."
---

# Introduction {-}

Reprex, a Dutch start-up enterprise formed to utilize open source software and open data, is looking for partners in an agile, open collaboration to win at least one of the three EU Datathon Prizes. We are looking for policy partners, academic partners and a consultancy partner. Our project is based on agile, open collaboration with three types of contributors.

With our competing prototypes we want to show that we have a research automation technology that can find open data, process it and validate it into high-quality business, policy or scientific indicators, and release it with daily refreshments in a modern API.

We are looking for institutions to challenge us with their data problems, and sponsors to increase our capacity. Over then next 5 months, we need to find a sustainable business model for a high-quality and open alternative to other public data programs.

## Mission Statement: Winning the EU Datathon 2021 Challenge {-}

- *To take part, you should propose the development of an application that links and uses open datasets.* - our [data curator team](https://music.dataobservatory.eu/#contributors)

- *Your application ... is also expected to find suitable new approaches and solutions to help Europe achieve important goals set by the European Commission through the use of open data.*” - this application is developed by our [technology contributors](https://greendeal.dataobservatory.eu/#contributors)

- *Your application should showcase opportunities for concrete business models or social enterprises.* - our [service development team](https://economy.dataobservatory.eu/#contributors) is working to make this happen!

- We use open source software and open data. The applications are hosted on the cloud resources of [Reprex](#reprex), an early-stage technology startup currently building a viable, open-source, open-data business model to create reproducible research products.

- We are working together with experts in the domain as curators (check out our guidelines if you want to join: [Data Curators: Get Inspired!](https://curators.dataobservatory.eu/data-curators.html)).

- Our development team works on an open collaboration basis. Our indicator R packages, and our services are developed together with [rOpenGov](https://music.dataobservatory.eu/author/ropengov/).

We want to win an [EU Datathon prize](https://op.europa.eu/en/web/eudatathon) by processing the vast, already-available governmental and scientific open data made usable for policy-makers, scientific researchers, and business researcher end-users. We aim to win at least one first prize in the EU Datathon 2021. We are contesting **all three** challenges, which are related to the EU’s official strategic policies for the coming decade.

## Challenge 1: A European Green Deal {-}

The Challenge [A European Green Deal](https://ec.europa.eu/info/strategy/priorities-2019-2024/european-green-deal_en), with a particular focus on the [The European Climate Pact](https://ec.europa.eu/commission/presscorner/detail/en/ip_20_2323), the [Organic Action Plan](https://ec.europa.eu/info/food-farming-fisheries/farming/organic-farming/organic-action-plan_en), and the [New European Bauhaus](https://ec.europa.eu/commission/presscorner/detail/en/IP_21_111), i.e., mitigation strategies. Our [Green Deal Data Observatory](http://greendeal.dataobservatory.eu/) is a modern reimagination of existing ‘data observatories’; currently, there are over 70 permanent international data collection and dissemination points. One of our objectives is to understand why the dozens of the EU’s observatories do not use open data and reproducible research. We want to show that open governmental data, open science, and reproducible research can lead to a higher quality and faster data ecosystem that fosters growth for policy, business, and academic data users.

```{r gdo, fig.out="85%", fig.align="center", echo=FALSE}
knitr::include_graphics(file.path(
  "plots", "screenshots", "greendeal_and_zenodo.png"
))
```

Please, follow us on social media, it really helps us finding new users and showing that we are able to grow our ecosystem: the [Green Deal Data Observatory on Linkedin](https://www.linkedin.com/company/78556699) and the [Green Deal Data Observatory on Twitter](https://twitter.com/GreenDealObs).

## Challenge 2: An economy that works for people {-}

Challenge 2: [An economy that works for people](https://ec.europa.eu/info/strategy/priorities-2019-2024/economy-works-people_en#:~:text=Individuals%20and%20businesses%20in%20the,needs%20of%20the%20EU's%20citizens.), with a particular focus on the [Single market strategy](https://ec.europa.eu/info/strategy/priorities-2019-2024/economy-works-people/internal-market_en), and particular attention to the strategy’s goals of 1. Modernising our standards system, 2. Consolidating Europe’s intellectual property framework, and 3. Enabling the balanced development of the collaborative economy strategic goals.

Big data and automation create new inequalities and injustices and have the potential to create a jobless growth economy. Our [Economy Data Observatory](https://economy.dataobservatory.eu/) is a fully automated, open source, open data observatory that produces new indicators from open data sources and experimental big data sources, with authoritative copies and a modern API.

Our observatory monitors the European economy to protect consumers and small companies from unfair competition, both from data and knowledge monopolization and robotization. We take a critical Small and Medium-Sized Enterprises (SME)-, intellectual property, and competition policy point of view of automation, robotization, and the AI revolution on the service-oriented European social market economy.

```{r edo-zenodo, fig.out="85%", fig.align="center", echo=FALSE}
knitr::include_graphics(file.path(
  "plots", "screenshots", "edo_and_zenodo.png"
))
```

Please, follow us on social media, it really helps us finding new users and showing that we are able to grow our ecosystem: the [Economy Data Observatory on Linkedin](https://www.linkedin.com/company/78562153) and the [Economy Data Observatory on Twitter](https://twitter.com/GreenDealObs) and join our [curatorial team](#economy).

## Challenge 3: A Europe fit for the digital age {-}

```{r dmo, fig.out="85%", fig.align="center", echo=FALSE}
knitr::include_graphics(file.path(
  "plots", "open_source_slide.png"
))
```


Challenge 3: [A Europe fit for the digital age](https://ec.europa.eu/info/strategy/priorities-2019-2024/europe-fit-digital-age_en), with a particular focus [Artificial Intelligence](https://ec.europa.eu/info/strategy/priorities-2019-2024/europe-fit-digital-age/excellence-trust-artificial-intelligence_en), the [European Data Strategy](https://ec.europa.eu/info/strategy/priorities-2019-2024/europe-fit-digital-age/european-data-strategy_en), the [Digital Services Act](https://ec.europa.eu/info/strategy/priorities-2019-2024/europe-fit-digital-age/digital-services-act-ensuring-safe-and-accountable-online-environment_en), [Digital Skills](https://digital-strategy.ec.europa.eu/en/policies/digital-skills-and-jobs) and [Connectivity](https://digital-strategy.ec.europa.eu/en/policies/connectivity). 

The [Digital Music Observatory](https://music.dataobservatory.eu/) is a fully automated, open source, open data observatory that creates public datasets to provide a comprehensive view of the European music industry. It provides high-quality and timely indicators in all four pillars of the planned official European Music Observatory as a modern, open source and largely open data-based, automated, API-supported alternative solution for this planned observatory. The insight and methodologies we are refining in the DMO are applicable and transferable to about 60 other data observatories funded by the EU which do not currently employ governmental or scientific open data.

```{r edo, fig.out="85%", fig.align="center", echo=FALSE}
knitr::include_graphics(file.path(
  "plots", "screenshots", "dmo_and_zenodo.png"
))
```


Please, follow us on social media, it really helps us finding new users and showing that we are able to grow our ecosystem: the [Digital Music Data Observatory on Linkedin](https://www.linkedin.com/company/reprexbv/) and the [Digital Music Data Observatory on Twitter](https://twitter.com/dataandlyrics).

## Problem Statement {-}

The EU has an 18-year-old open data regime and it makes public taxpayer-funded data in the values of tens of billions of euros per year; the Eurostat program alone handles 20,000 international data products, including at least 5,000 pan-European environmental indicators.

As open science principles gain increased acceptance, scientific researchers are making hundreds of thousands of valuable datasets public and available for replication every year.

The EU, the OECD, and UN institutions run around 100 data collection programs, so-called ‘data observatories’ that more or less avoid touching this data, and buy proprietary data instead. Annually, each observatory spends between 50 thousand and 3 million EUR on collecting untidy and proprietary data of inconsistent quality, while never even considering open data.


```{r problem-obs-collage, fig.out="85%", fig.align="center", echo=FALSE}
knitr::include_graphics(file.path(
  "plots", "screenshots", "observatory_collage_16x9_800.png"
))
```


The problem with the current EU data strategy is that while it produces enormous quantities of valuable open data, in the absence of common basic data science and documentation principles, it seems often cheaper to create new data than to put the existing open data into shape.

This is an absolute waste of resources and efforts. With a few R packages and our deep understanding of advanced data science techniques, we can create valuable datasets from unprocessed open data. In most domains, we are able to repurpose data originally created for other purposes at a historical cost of several billions of euros, converting these unused data assets into valuable datasets that can replace tens of millions’ worth of proprietary data.

What we want to achieve with this project – and we believe such an accomplishment would merit one of the first prizes - is to add value to a significant portion of pre-existing EU open data (for example, available on [data.europa.eu/data](https://data.europa.eu/data/)) by re-processing and integrating them into a modern, tidy database with an API access, and to find a business model that emphasises a triangular use of data in 1. business, 2. science and 3. policy-making. Our mission is to modernize the concept of `data observatories.`

## Our solution {-}

We are creating an agile open collaboration network with 
    - [data curators](#data-curators) who can tell us what data is needed in climate change policies, climate politics, computational antitrust, innovation economics, economic and enviromental impact assessments, data governance impact assessments, trustworhty AI impact assessement;
    - [open data developers](#tech) to deliver open data in an automated, data-as-service application (see our [technical skillset needs](#service-flow)); 
    - [service designers](#service) who help us build a reliable service and business case.

Our Product/Market Fit was validated in the world’s 2nd ranked university-backed incubator program, the [Yes!Delft AI Validation Lab](https://music.dataobservatory.eu/post/2020-09-25-yesdelft-validation/). We are currently developing this project with the help of the [JUMP European Music Market Accelerator](https://www.jumpmusic.eu/fellow2021/automated-music-observatory/) program.
